---
title: "Economic Value"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this analysis, we will assess the economic value of the crops grown in this
home garden case study. All the economic data for this analysis was collected 
from 2020 even though the crops were harvested from 2011-2013. However, to make 
this analysis as relevant as possible, more current economic data will be used. 
All prices were found online from supermarkets including Wegmans, WalmartFood, 
and Amazon Pantry. Each of these crops were priced by pound for both organic and 
conventional prices. For those crops that could not be found in organic and 
conventional varieties, a average of all the organic to conventional crop ratios 
was taken and used to estimate the price of whichever crop type we did not have.
Because the crops were organically grown, organic prices were used. Let's begin! 

First, we have to upload our packages. 

```{r package upload, message = FALSE}
library(dplyr)
library(ggplot2)
```

With our packages uploaded, we can now upload and join our data. 

```{r data upload, message = FALSE}
garden.data.raw <- read.csv("~/Documents/Garden Data Analysis/Garden Data Full.csv") %>%
  mutate(weight_lbs = weight_oz * 0.0625)
crop_key <- read.csv("~/Documents/Garden Data Analysis/crop_key_species_money.csv")

garden.data.species <- inner_join(garden.data.raw, crop_key, by = NULL, 
                                  copy = FALSE, suffix = c(".garden.data.raw",
                                                           ".crop_key"))
```

The data needed for analysis is now uploaded, but we need to organize it in a 
way that makes it usable. We need to group all the species data together and 
calculate the total pounds of each species per year. We can also calculate the 
price of the yield with the economic data we have. 

```{r $/lbs}
garden.data.species$price_pound_conventional <- as.numeric(garden.data.species$price_pound_conventional)

total_lbs_species_yield <- garden.data.species %>%
  mutate(yield_O = weight_lbs * price_pound_organic) %>%
  mutate(yield_conven = weight_lbs * price_pound_conventional) %>%
  group_by(year, species, family)

lbs_species_yield <- total_lbs_species_yield %>%
  summarize(total_lbs_con = sum(yield_conven), total_lbs_O = sum(yield_O)) 
```

With the economic yield calculated and summarized by year and species, we can 
begin graphing this data. We will start by looking at the economic yield by year 
for the conventional crop prices. 

```{r violin conventional}
lbs_species_yield$year <- as.character.Date(lbs_species_yield$year)
ggplot(lbs_species_yield, aes(x = year, y = total_lbs_con)) + 
  geom_violin() + scale_y_log10()
```

We can also look to see how much value the crops hold as organic crops. 

```{r violin organic}
lbs_species_yield$year <- as.character.Date(lbs_species_yield$year)
ggplot(lbs_species_yield, aes(x = year, y = total_lbs_O)) + 
  geom_violin() + scale_y_log10()
```

We can also compare the organic versus the conventional prices. To visualize 
this, we will make a data table. 

```{r yearly_yield}
total_yearly_yield <- lbs_species_yield %>%
  group_by(year) %>%
  summarize(yearly_yield_con = sum(total_lbs_con), 
            yearly_yield_O = sum(total_lbs_O)) 

total_yearly_yield
```

Based on our data table, 2013 actually had the highest economic value of the 
crops, valued at $2,237.89 for organic produce! This home garden was sure 
worth a lot! 

We can also visualize this by family by year. We will do this with the organic 
yields as the crops were grown organically.

```{r yield_family}
ggplot(lbs_species_yield, aes(x = year, y = total_lbs_O, fill = family, 
                                    color = family)) + geom_col()
```

The family Solanaceae seems to have the greatest economic yield. We can further 
investigate this to see which crop species had the highest economic value. 

```{r species_yield}
solanaceae <- lbs_species_yield %>%
  filter(family == "Solanaceae")

ggplot(solanaceae, aes(x = year, y = total_lbs_O, fill = species, 
                       color = species)) + geom_col()
```

Based on this data, *Solanum lycopersicum*, which includes all the tomato 
varieties, was the most economically productive in the garden! 

