---
title: 'Productivity Part 1: lbs/acre'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of this analysis is to understand how productive our home garden case 
study was. The home garden data went from 2011-2013 in a 600 square foot garden. 
However, we estimate that about 325 square foot of that garden was actually in 
production. For this first step of understanding productivity, we are analyzing 
the harvest by pounds per acre. Let's begin! 

First, we will upload our packages. 

```{r package upload, message = FALSE}
library(dplyr)
library(ggplot2)
```

Now, we have to upload our datasets and prepare them for analysis. 

```{r data upload, message = FALSE}
garden.data.raw <- read.csv("~/Documents/Garden Data Analysis/Garden Data Full.csv") %>%
  mutate(weight_lbs = weight_oz * 0.0625)
crop_key_species <- read.csv("~/Documents/Garden Data Analysis/crop_key_species.csv")
garden.data.species <- inner_join(garden.data.raw, crop_key_species, by = NULL,
                                  copy = FALSE, suffix = c(".garden.data.raw",
                                                           ".crop_key_species"))
```

With the data set now available and ready, we want to create a few easy conversion
variables to use. We want everything in terms of acres, so we need to convert the
amount of land in square feet to acres. We will do this for both the total land 
and the productive land. 

```{r variable assigning}
a <- 600 * 2.29568411 * 10^-5
b <- 325 * 2.29568411 * 10^-5
```

Now we can calculate lbs/acre. To do this, we will add an additional column to 
our data using mutate(). 

```{r adding acre data}
acre_all_year <- garden.data.species %>%
  mutate(lbs_acre = weight_lbs * a) %>%
  mutate(product_acre = weight_lbs * b)
```

We now have all our necessary data in one place, but our 'year' column is classified
as an integer when we need it to be a character. Luckily, this change is fairly 
easy to make. 

```{r year to character}
acre_all_year$year <- as.character.Date(acre_all_year$year)
```

With these columns added, one for the total area of the garden and one for the 
productive part of the garden, and the year classified properly, we can make 
violin plots to show productivity in lbs/acre per year for total area. 

```{r total acre violin}
ggplot(acre_all_year, aes(x = year, y = lbs_acre)) + geom_violin() + 
  scale_y_log10()
```

We can also make one to show just the productive part of the garden. 

```{r productive acre violin}
ggplot(acre_all_year, aes(x = year, y = product_acre)) + geom_violin() + 
  scale_y_log10()
```

With two violin plots showing our yearly productivity, we can also look at the 
productivity by family. 

```{r total acre family}
ggplot(acre_all_year, aes(x = year, y = lbs_acre, fill = family, 
                          color = family)) + geom_col()
```

We can see that by weight, family Solanaceae was the most productive! However, 
which crop of family Solanaceae was the more productive? By isolating the family, 
we can find out! 

```{r solanaceae total acre}
solanaceae <- acre_all_year %>%
  filter(family == "Solanaceae")

ggplot(solanaceae, aes(x = year, y = lbs_acre, fill = species, 
                       color = species)) + geom_col()
```

Based on this graph, species *Solanum lycopersicum* was the most productive 
species in the garden. This species contains tomatoes, of which the case study 
grew several varieties. Tomatoes were the most productive crop by weight probably
because of their high water content, making them quite heavy. Based on this 
analysis,tomatoes were the most productive crop by lbs/acre. 



