---
title: "Shannon Index Calculations"
author: "Zoe"
date: "7/1/2020"
output: html_document
---

In order to assess biodiversity of an ecosystem, the Shannon Wiener Biodiversity
Index is often used by ecologists.  This simple formula takes proportions of the 
individual plant abundance to the total plant abundance.  With this case study, 
we used the weight of the harvested plants to calculate diversity.  This limits
our study as some plants have more water content than other and will therefore
be heavier.  Though this is a limited way of examining the diversity of the 
gardens, it is th best way for us to calculate the Shannon index with the data 
collected.  

As always, we first needed to start our packages and upload our data. We also
will join the data we have to make it easier to use.  

```{r}
library(dplyr)
library(workflowr)

garden.data.raw <- read.csv("~/Documents/Garden Data Analysis/Garden Data Full.csv") %>%
  mutate(weight_llbs = weight_oz * 0.0625)
crop.families <- read.csv("~/Documents/Garden Data Analysis/crop_key.csv")

garden.data.families <- full_join(garden.data.raw, crop.families, by = NULL,
                                  copy = FALSE, suffix = c(".garden.data.raw",
                                                           ".crop.families"))
garden.data.families$count <- as.numeric(garden.data.families$count)
```

Once the data is prepared, we began isolating the data by year to make it easier
to calculate the Shannon Index by year. For this summary, we will look at 
2011 only. 

```{r}
crops.2011 <- garden.data.families %>%
  filter(year == 2011)
```

With the year isolated from the raw dataset, we also wanted to isolate the crops
and the weight_llbs
columns. To do this, we created a new data frame with only these columns. 

```{r}
factor.2011.crop <- crops.2011$crop
factor.2011.weight_llbs <- crops.2011$weight_llbs
crop.2011.weights <- data.frame(factor.2011.crop, factor.2011.weight_llbs)
```

With the necessary data now in an isolated format, we then wanted to sum the 
weight_llbs for each crop species. Another data frame was generated with the 
total_llbs by crop species.  

```{r}
total.2011.weights <- crop.2011.weights %>%
  group_by(factor.2011.crop) %>%
  summarize(total_llbs = sum(factor.2011.weight_llbs))
```

With the total weights for 2011 calculated, we wanted to get the total weight
of all the crops together. This would allow us to calculate the proportions of 
specific crop weight to total crop weight.  
```{r}
total.2011.weight.all.crop <- sum(factor.2011.weight_llbs)
total.2011.weight.all.crop
```

With these calculations finished, we could begin using the Shannon Index. The
formula for the Shannon Index is H' = -sum[pi(ln(pi))] where pi is the proportion
of weight from individual crops by the total weight of the crops. 

```{r}
p.2011 <- total.2011.weights %>%
  mutate(p = total_llbs / total.2011.weight.all.crop) %>%
  mutate(ln = log(p)) %>%
  mutate(pln = ln * p)
H.2011 <- sum(p.2011$pln) * -1
H.2011
H.2012 <- 3.084806
H.2013 <- 2.901443
```

The Shannon Index has now been calculated! Each value was put into a data table
to best see the differences between the years.  
```{r}
years <- c("2011", "2012", "2013")
shannon_index <- c(H.2011, H.2012, H.2013)
shannon_index_all_years <- data.frame(years, shannon_index)
shannon_index_all_years
```

High Shannon Index values like this indicates high diversity of the garden over 
all three years. On the next page, the species evenness will be calculated. 
