---
title: "Species Richness"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The goal of this analysis is to calculate the species richness from our case study
garden. Species richness is a measure of biodiversity that states how many different
species are found in an ecosystem.  

To start, we made sure to load the necessary packages onto RStudio.  

``` {r, results = 'hide'}
library(dplyr)
library(ggplot2)
library(workflowr)
```

Once the necessary packages were loaded, we needed to access our data.  There were
two datasets used in this analysis, including the original harvest log, known as 
"Garden Data Full.csv" and a list of the crop families, listed as "crop_key.csv". 
When we uploaded this data, we renamed it "garden.data.raw" and "crop.families" 
respectively. Additionally, we wanted to add a column onto the garden.data.raw that 
converted the weight in ounces to pounds. 

```{r}
garden.data.raw <- read.csv("~/Documents/Garden Data Analysis/Garden Data Full.csv") %>%
  mutate(weight_llbs = weight_oz * 0.0625)
crop.families <- read.csv("~/Documents/Garden Data Analysis/crop_key_species.csv")
```

We now have both data sets in our environment, but we want to join these two into
one, full data frame.  To do this, we used the full_join function 

```{r, results = 'hide'}
garden.data.families <- full_join(garden.data.raw, crop.families, by = NULL, 
                                  copy = FALSE, suffix = c(".garden.data.raw", 
                                                           ".crop.families"))
```

The data sets have now been joined into one large data set that will be easier 
to work with.  The only issue with this data set is that it is using the "count"
variable as a character when we want it to be a numeric variable.  Essentially, 
the calculations we want to use with count will not work unless we change the 
class type of this variable.  

```{r}
garden.data.families$count <- as.numeric(garden.data.families$count)
str(garden.data.families)
```

The variable "count" is now numeric, which will make calculations easier.  Now
we can start calculations for species richness.  We want to know the species 
richness by year, so first we should isolate each year.  

```{r}
crops.2011 <- garden.data.families %>%
  group_by(year) %>%
  filter(year == 2011)

crops.2012 <- garden.data.families %>%
  group_by(year) %>%
  filter(year == 2012)

crops.2013 <- garden.data.families %>%
  group_by(year) %>%
  filter(year == 2013)
```

The years are now isolated so now we have to learn how many crops were planted
each year.  To do this, the years were made into factors and then summarized by 
the species. We will use 2011 as an example.  

```{r}
factor.2011 <- factor(crops.2011$species)
summary(factor.2011)
```

Based on this analysis, 2011 had 26 crop varieties.  By doing the same analysis 
for each year, 2012 had 44 crop varieties and 2013 had 39 crop varieties.  Knowing
the values is helpful, but being able to visualize it is much more important.  
In order to visualize this data, a separate data frame was formed that would then
be used to make a graph. 

```{r}
year <- c("2011", "2012", "2013")
richness <- c("26", "25", "25")

richness.by.year <- data.frame(year, richness)
richness.by.year
```

Finally, we have all the data in the correct format to make a graph.  For this data,
using a bar graph will be most useful for seeing the richness by year. 

```{r}
ggplot(richness.by.year, aes(x = year, y = richness)) + geom_col()
```



