---
title: "Species Evenness"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

With these calculations, we hope to calulate the species evenness over all three
years of this case study garden. We already organized the data in the other two
calculations, so to see the coding to organize the data, look to 
[Species Richness](Species-Richness-Rmark.html) or [Shannon Index](Shannon-Index.html). Below are the data we need for this analysis. 

```{r}
library(dplyr)

richness <- c("26", "25", "25")
richness <- as.numeric(richness)
years <- c("2011", "2012", "2013")
H.2011 <- 3.042144
H.2012 <- 3.084806
H.2013 <- 2.901443
shannon_index <- c(H.2011, H.2012, H.2013)
```

As with the Shannon Index calculations, these will be done by weight in llbs of
the harvested crops. This is limited because harvested produce all have different
weights. However, with the data available, this method will be the most accurate
to the true species evenness of the garden. 

As done on the other pages, we will keep the data isolated by year. In order 
calculate evenness, we will use the following equation: 
J = H' / lnS, where J is the species evenness, H' is the Shannon Index calulated 
[here](Shannon-Index.html), and S is the species richness.  

Luckily enough, we have already calculated the species richness and Shannon Index!
All we have to do is plug these values into the equation. For refresher sake, 
here is a table with the species richness and Shannon Index values. 

```{r}
J.prep <- data.frame(years, richness, shannon_index)
J.prep
```

Now, we plug the values into the equation!

```{r}
species_evenness <- J.prep %>%
  mutate(species.evenness = shannon_index / log(richness))
species_evenness
```
Now the species evenness has been calculated! As seen in the table, the species
evenness across the years is fairly similar. Because the digits are fairly close
to one, this habitat is fairly even. If the values were 0, the garden would be 
completely uneven. Happily, according to weights of the harvested crops, this
garden was species even. 

